---
title: "Regression Model: "
author: "Anna Vannucci"
date: "6/3/2022"
output: html_document
---

resources: 
https://cran.r-project.org/web/packages/brms/vignettes/brms_multivariate.html
http://mjskay.github.io/tidybayes/articles/tidy-brms.html

Load libraries:
```{r}
library(tidyverse)
library(psych)
library(viridis)
library(brms)
library(loo)
library(tidybayes)
library(bayesplot)
library(posterior)
library(cowplot)
library(purrr)
library(modelr)
library(caret)
library(ggmcmc)
library(ggthemes)
library(ggridges)

# help stan run faster
#rstan_options(auto_write = T)
#options(mc.cores = parallel::detectCores())

set.seed(122) # for reproducibility
```

Load dataframes:
```{r}
# read in cleaned dataframe
df <- read.csv("../data_clean/df_wide.csv", stringsAsFactors = F)

# read in dataframe with site labels
dfsite <- read.csv("../data_raw/abcd_lt01.csv", stringsAsFactors = F) %>% 
  filter(eventname == "baseline_year_1_arm_1") %>% 
  dplyr::select(subjectkey, site_id_l)

# add site variable to cleaned dataframe
df <- df %>% 
  left_join(dfsite, by = "subjectkey") %>% 
  drop_na()
```

Data cleaning:
```{r}
df_clean <- df %>% 
  dplyr::select(altID, subjectkey, site_id_l, age, sex, f1_parpsych, f2_ses_adi_ncrime, 
                f3_p2lacksupp, f4_p1lacksupp, f5_famconf_cr, f6_parsu_biosep, f7_angarg_pr, 
                f8_physemoagg_pr, f9_trauma, f10_neglos, cbcl_int_t, cbcl_ext_t, cardsort_uc, 
                flanker_uc) %>% 
  dplyr::rename(site = site_id_l) %>% 
  dplyr::mutate(site = as.factor(site),
                age_c = age - mean(age),
                sex_d = as.factor(ifelse(sex == "M", -0.5, 0.5)),
                f1_parpsych_c = f1_parpsych - mean(f1_parpsych), 
                f2_ses_adi_ncrime_c = f2_ses_adi_ncrime - mean(f2_ses_adi_ncrime), 
                f3_p2lacksupp_c = f3_p2lacksupp - mean(f3_p2lacksupp), 
                f4_p1lacksupp_c = f4_p1lacksupp - mean(f4_p1lacksupp), 
                f5_famconf_cr_c = f5_famconf_cr - mean(f5_famconf_cr), 
                f6_parsu_biosep_c = f6_parsu_biosep - mean(f6_parsu_biosep), 
                f7_angarg_pr_c = f7_angarg_pr - mean(f7_angarg_pr), 
                f8_physemoagg_pr_c = f8_physemoagg_pr - mean(f8_physemoagg_pr), 
                f9_trauma_c = f9_trauma - mean(f9_trauma), 
                f10_neglos_c = f10_neglos - mean(f10_neglos),
                cbclint_std = scale(cbcl_int_t),
                cbclext_std = scale(cbcl_ext_t),
                cardsort_std = scale(cardsort_uc),
                flanker_std = scale(flanker_uc))

write.csv(df_clean, "../data_clean/df_mvmlm.csv", row.names = F)
```

Split dataframe into train and validation (test) sets:
```{r}
trainIndex <- createDataPartition(df_clean$site, 
                                  p = .7,
                                  list = FALSE,
                                  times = 1)
train <- df_clean[ trainIndex,]
valid <- df_clean[-trainIndex,]
```

Sample characteristics of train and test sets:
```{r}
describe(train)
describe(valid)
```


***********************
***********************
**** TRAINING SET *****
***********************
***********************

# Data cleaning
```{r}
df_train <- train %>% 
  dplyr::mutate(site = as.factor(site),
                age_c = age - mean(age),
                sex_d = as.factor(ifelse(sex == "M", -0.5, 0.5)),
                f1_parpsych_c = f1_parpsych - mean(f1_parpsych), 
                f2_ses_adi_ncrime_c = f2_ses_adi_ncrime - mean(f2_ses_adi_ncrime), 
                f3_p2lacksupp_c = f3_p2lacksupp - mean(f3_p2lacksupp), 
                f4_p1lacksupp_c = f4_p1lacksupp - mean(f4_p1lacksupp), 
                f5_famconf_cr_c = f5_famconf_cr - mean(f5_famconf_cr), 
                f6_parsu_biosep_c = f6_parsu_biosep - mean(f6_parsu_biosep), 
                f7_angarg_pr_c = f7_angarg_pr - mean(f7_angarg_pr), 
                f8_physemoagg_pr_c = f8_physemoagg_pr - mean(f8_physemoagg_pr), 
                f9_trauma_c = f9_trauma - mean(f9_trauma), 
                f10_neglos_c = f10_neglos - mean(f10_neglos),
                cbclint_std = scale(cbcl_int_t),
                cbclext_std = scale(cbcl_ext_t),
                cardsort_std = scale(cardsort_uc),
                flanker_std = scale(flanker_uc))

write.csv(df_train, "../data_clean/df_train.csv", row.names = F)
```

```{r}
df_valid <- valid %>% 
  dplyr::mutate(site = as.factor(site),
                age_c = age - mean(age),
                sex_d = as.factor(ifelse(sex == "M", -0.5, 0.5)),
                f1_parpsych_c = f1_parpsych - mean(f1_parpsych), 
                f2_ses_adi_ncrime_c = f2_ses_adi_ncrime - mean(f2_ses_adi_ncrime), 
                f3_p2lacksupp_c = f3_p2lacksupp - mean(f3_p2lacksupp), 
                f4_p1lacksupp_c = f4_p1lacksupp - mean(f4_p1lacksupp), 
                f5_famconf_cr_c = f5_famconf_cr - mean(f5_famconf_cr), 
                f6_parsu_biosep_c = f6_parsu_biosep - mean(f6_parsu_biosep), 
                f7_angarg_pr_c = f7_angarg_pr - mean(f7_angarg_pr), 
                f8_physemoagg_pr_c = f8_physemoagg_pr - mean(f8_physemoagg_pr), 
                f9_trauma_c = f9_trauma - mean(f9_trauma), 
                f10_neglos_c = f10_neglos - mean(f10_neglos),
                cbclint_std = scale(cbcl_int_t),
                cbclext_std = scale(cbcl_ext_t),
                cardsort_std = scale(cardsort_uc),
                flanker_std = scale(flanker_uc))

write.csv(df_valid, "../data_clean/df_valid.csv", row.names = F)
```



# Multivariate multilevel models - TRAINING SET

*Internalizing and Externalizing Problems*

`Standardized  (beta) Estimates`
```{r}
# multivariate multilevel model equation
bform2_tr <- bf(mvbind(cbclint_std, cbclext_std) ~ scale(age_c) + sex_d + scale(f1_parpsych_c) +
               scale(f2_ses_adi_ncrime_c) + scale(f3_p2lacksupp_c) + scale(f4_p1lacksupp_c) + 
               scale(f5_famconf_cr_c) + scale(f6_parsu_biosep_c) + scale(f7_angarg_pr_c) + 
               scale(f8_physemoagg_pr_c) + scale(f9_trauma_c) + scale(f10_neglos_c) + 
               (1|a|site)) + # specify random intercepts
            set_rescor(TRUE) # specify a multivariate model; model residual correlations

# run mvmlm
fit2_tr <- brm(bform2_tr, data = df_train, chains = 2, cores = 2)
saveRDS(fit2_tr, "../results/mvmlm_train_cbcl_std.rds")
summary(fit2_tr)
```

`model fit`
- How much variation in the dependent variables can be explained by our model
- Use a Bayesian generalization of the R2 coefficient.
```{r}
fit2_tr_r2 <- bayes_R2(fit2_tr)
fit2_tr_r2
```

`prediction`
https://www.geeksforgeeks.org/the-validation-set-approach-in-r-programming/

*internalizing*
```{r}
fit2_tr <- readRDS("../results/mvmlm_train_cbcl_std.rds")
df_valid <- read.csv("../data_clean/df_valid.csv", stringsAsFactors = F)

# make predictions of new data based on model
predictions_int <- predict(fit2_tr, resp = "cbclintstd", newdata = df_valid)
saveRDS(predictions_int, file = "../results/mvmlm_predict_int.rds")
head(predictions_int)


# computing model performance metrics 
data.frame(R2 = R2(predictions_int, df_valid$cbclint_std), 
           RMSE = RMSE(predictions_int, df_valid$cbclint_std), 
           MAE = MAE(predictions_int, df_valid$cbclint_std))

# spearman correlation
rho_int = cor.test(predictions_int[,1], df_valid$cbclint_std)
rho_int
```

*externalizing*
```{r}
# make predictions of new data based on model
predictions_ext <- predict(fit2_tr, resp = "cbclextstd", newdata = df_valid)
saveRDS(predictions_ext, file = "../results/mvmlm_predict_ext.rds")
head(predictions_ext)

# computing model performance metrics 
data.frame(R2 = R2(predictions_ext, df_valid$cbclext_std), 
           RMSE = RMSE(predictions_ext, df_valid$cbclext_std), 
           MAE = MAE(predictions_ext, df_valid$cbclext_std))

rho_ext = cor.test(predictions_ext[,1], df_valid$cbclext_std)
rho_ext
```



*Cognitive Control*

`Standardized (beta) Estimates`
```{r}
# multivariate multilevel model equation
bform4_tr <- bf(mvbind(cardsort_std, flanker_std) ~ scale(age_c) + sex_d + scale(f1_parpsych_c) +
               scale(f2_ses_adi_ncrime_c) + scale(f3_p2lacksupp_c) + scale(f4_p1lacksupp_c) + 
               scale(f5_famconf_cr_c) + scale(f6_parsu_biosep_c) + scale(f7_angarg_pr_c) + 
               scale(f8_physemoagg_pr_c) + scale(f9_trauma_c) + scale(f10_neglos_c) + 
               (1|a|site)) + # specify random intercepts
            set_rescor(TRUE) # specify a multivariate model; model residual correlations

# run mvmlm
fit4_tr <- brm(bform4_tr, data = df_train, chains = 2, cores = 2)
saveRDS(fit4_tr, "../results/mvmlm_train_cogctrl_std.rds")
summary(fit4_tr)
```
`model fit`
- How much variation in the dependent variables can be explained by our model
- Use a Bayesian generalization of the R2 coefficient.
```{r}
fit4_tr_r2 <- bayes_R2(fit4_tr)
fit4_tr_r2
```

`prediction`
https://www.geeksforgeeks.org/the-validation-set-approach-in-r-programming/

*cognitive flexibility*
```{r}
fit4_tr <- readRDS("../results/mvmlm_train_cogctrl_std.rds")

# make predictions of new data based on model
predictions_cf <- predict(fit4_tr, resp = "cardsortstd", newdata = df_valid)
saveRDS(predictions_cf, "../results/mvmlm_predict_cogflx.rds")
head(predictions_cf)

# computing model performance metrics 
data.frame(R2 = R2(predictions_cf, df_valid$cardsort_std), 
           RMSE = RMSE(predictions_cf, df_valid$cardsort_std), 
           MAE = MAE(predictions_cf, df_valid$cardsort_std))

rho_cf = cor.test(predictions_cf[,1], df_valid$cardsort_std)
rho_cf
```

*inhibitory control*
```{r}

# make predictions of new data based on model
predictions_ic <- predict(fit4_tr, resp = "flankerstd", newdata = df_valid)
saveRDS(predictions_ic, "../results/mvmlm_predict_inhctrl.rds")
head(predictions_ic)

# computing model performance metrics 
data.frame(R2 = R2(predictions_ic, df_valid$flanker_std), 
           RMSE = RMSE(predictions_ic, df_valid$flanker_std), 
           MAE = MAE(predictions_ic, df_valid$flanker_std),
           r = cor(predictions_ic, df_valid$flanker_std))

rho_ic = cor.test(predictions_ic[,1], df_valid$flanker_std)
rho_ic
```




# Multivariate multilevel models - VALIDATION SET (for characterization of what effects replicated only)

*Internalizing and Externalizing Problems*

`Standardized  (beta) Estimates`
```{r}
# multivariate multilevel model equation
bform_cbcl_val <- bf(mvbind(cbclint_std, cbclext_std) ~ scale(age_c) + sex_d + scale(f1_parpsych_c) +
                       scale(f2_ses_adi_ncrime_c) + scale(f3_p2lacksupp_c) + scale(f4_p1lacksupp_c) +
                       scale(f5_famconf_cr_c) + scale(f6_parsu_biosep_c) + scale(f7_angarg_pr_c) +
                       scale(f8_physemoagg_pr_c) + scale(f9_trauma_c) + scale(f10_neglos_c) + 
                       (1|a|site)) + # specify random intercepts
                       set_rescor(TRUE) # specify a multivariate model; model residual correlations

# run mvmlm
fit_cbcl_val <- brm(bform_cbcl_val, data = df_valid, chains = 2, cores = 2)
saveRDS(fit_cbcl_val, "../results/mvmlm_valid_cbcl_std.rds")
summary(fit_cbcl_val)
```

*Cognitive Control*

`Standardized (beta) Estimates`
```{r}
# multivariate multilevel model equation
bform_cc_val <- bf(mvbind(cardsort_std, flanker_std) ~ scale(age_c) + sex_d + scale(f1_parpsych_c) +
                     scale(f2_ses_adi_ncrime_c) + scale(f3_p2lacksupp_c) + scale(f4_p1lacksupp_c) + 
                     scale(f5_famconf_cr_c) + scale(f6_parsu_biosep_c) + scale(f7_angarg_pr_c) + 
                     scale(f8_physemoagg_pr_c) + scale(f9_trauma_c) + scale(f10_neglos_c) + 
                     (1|a|site)) + # specify random intercepts
                    set_rescor(TRUE) # specify a multivariate model; model residual correlations

# run mvmlm
fit_cc_val <- brm(bform_cc_val, data = df_valid, chains = 2, cores = 2)
saveRDS(fit_cc_val, "../results/mvmlm_valid_cogctrl_std.rds")
summary(fit_cc_val)
```



**plots**

```{r}
ic_pred <- as.numeric(predictions_ic[,1])
df_valid$pred_ic <- ic_pred

df_valid %>% 
  ggplot(aes(x = pred_ic, y = flanker_std)) +
  geom_jitter()
```

