---
title: "Plots of Results from Multivariate Multilevel Models"
author: "Anna Vannucci"
date: "6/5/2022"
output: html_document
---

Load libraries:
```{r}
library(tidyverse)
library(psych)
library(viridis)
library(brms)
library(loo)
library(tidybayes)
library(bayesplot)
library(posterior)
library(cowplot)
library(purrr)
library(modelr)
library(caret)
library(ggmcmc)
library(ggthemes)
library(ggridges)
library(rstan)
library(gghalves)

# help stan run faster
#rstan_options(auto_write = T)
#options(mc.cores = parallel::detectCores())

set.seed(122) # for reproducibility
```


Upload dataframes and model results:
```{r}
# dataframes
df_train <- read.csv("../data_clean/df_train.csv", stringsAsFactors = F)
df_valid <- read.csv("../data_clean/df_valid.csv", stringsAsFactors = F)

# mvmlm model training in the discovery set
m_cbcl_tr <- readRDS("../results/mvmlm_train_cbcl_std.rds")
m_cc_tr <- readRDS("../results/mvmlm_train_cogctrl_std.rds")

# prediction models
m_pred_int <- readRDS("../results/mvmlm_predict_int.rds")
m_pred_ext <- readRDS("../results/mvmlm_predict_ext.rds")
m_pred_cf <- readRDS("../results/mvmlm_predict_cogflx.rds")
m_pred_ic <- readRDS("../results/mvmlm_predict_inhctrl.rds")

# mvmlm models in the replication/validation set
m_cbcl_val <- readRDS("../results/mvmlm_valid_cbcl_std.rds")
m_cc_val <- readRDS("../results/mvmlm_valid_cogctrl_std.rds")
```


Data cleaning:
```{r}
# create numeric vectors of child behavior predicted scores
pred_int <- as.numeric(m_pred_int[,1]) # internalizing problems
pred_ext <- as.numeric(m_pred_ext[,1]) # externalizing problems
pred_cf <- as.numeric(m_pred_cf[,1]) # cognitive flexibility
pred_ic <- as.numeric(m_pred_ic[,1]) # inhibitory control

# predicted scores to validation dataset
df_valid$pred_int <- pred_int
df_valid$pred_ext <- pred_ext
df_valid$pred_cf <- pred_cf
df_valid$pred_ic <- pred_ic
```


# Results of MVMLM: posterior estimates of association between ELA and child behavior 

a) view a list of variables that can be extracted
```{r}
get_variables(m_cbcl_tr)
```
`internalizing/externalizing problems`

a) prepare data for plotting
```{r}
# extract posteriors from mvmlm model for int/ext problems
post_cbcl_long <- m_cbcl_tr %>% 
  spread_draws(b_cbclintstd_Intercept, b_cbclintstd_scalef1_parpsych_c,
               b_cbclintstd_scalef2_ses_adi_ncrime_c, b_cbclintstd_scalef3_p2lacksupp_c, 
               b_cbclintstd_scalef4_p1lacksupp_c, b_cbclintstd_scalef5_famconf_cr_c,
               b_cbclintstd_scalef6_parsu_biosep_c, b_cbclintstd_scalef7_angarg_pr_c, 
               b_cbclintstd_scalef8_physemoagg_pr_c, b_cbclintstd_scalef9_trauma_c,
               b_cbclintstd_scalef10_neglos_c,
               b_cbclextstd_Intercept, b_cbclextstd_scalef1_parpsych_c,
               b_cbclextstd_scalef2_ses_adi_ncrime_c, b_cbclextstd_scalef3_p2lacksupp_c, 
               b_cbclextstd_scalef4_p1lacksupp_c, b_cbclextstd_scalef5_famconf_cr_c,
               b_cbclextstd_scalef6_parsu_biosep_c, b_cbclextstd_scalef7_angarg_pr_c, 
               b_cbclextstd_scalef8_physemoagg_pr_c, b_cbclextstd_scalef9_trauma_c,
               b_cbclextstd_scalef10_neglos_c) %>% 
  mutate(f1parpsych_int = b_cbclintstd_Intercept + b_cbclintstd_scalef1_parpsych_c,
         f2ses_int = b_cbclintstd_Intercept + b_cbclintstd_scalef2_ses_adi_ncrime_c,
         f3p2nosupp_int = b_cbclintstd_Intercept + b_cbclintstd_scalef3_p2lacksupp_c,
         f4p1nosupp_int = b_cbclintstd_Intercept + b_cbclintstd_scalef4_p1lacksupp_c,
         f5famconf_int = b_cbclintstd_Intercept + b_cbclintstd_scalef5_famconf_cr_c,
         f6psusep_int = b_cbclintstd_Intercept + b_cbclintstd_scalef6_parsu_biosep_c,
         f7angarg_int = b_cbclintstd_Intercept + b_cbclintstd_scalef7_angarg_pr_c,
         f8famagg_int = b_cbclintstd_Intercept + b_cbclintstd_scalef8_physemoagg_pr_c,
         f9trauma_int = b_cbclintstd_Intercept + b_cbclintstd_scalef9_trauma_c,
         f10parlos_int = b_cbclintstd_Intercept + b_cbclintstd_scalef10_neglos_c,
         f1parpsych_ext = b_cbclextstd_Intercept + b_cbclextstd_scalef1_parpsych_c,
         f2ses_ext = b_cbclextstd_Intercept + b_cbclextstd_scalef2_ses_adi_ncrime_c,
         f3p2nosupp_ext = b_cbclextstd_Intercept + b_cbclextstd_scalef3_p2lacksupp_c,
         f4p1nosupp_ext = b_cbclextstd_Intercept + b_cbclextstd_scalef4_p1lacksupp_c,
         f5famconf_ext = b_cbclextstd_Intercept + b_cbclextstd_scalef5_famconf_cr_c,
         f6psusep_ext = b_cbclextstd_Intercept + b_cbclextstd_scalef6_parsu_biosep_c,
         f7angarg_ext = b_cbclextstd_Intercept + b_cbclextstd_scalef7_angarg_pr_c,
         f8famagg_ext = b_cbclextstd_Intercept + b_cbclextstd_scalef8_physemoagg_pr_c,
         f9trauma_ext = b_cbclextstd_Intercept + b_cbclextstd_scalef9_trauma_c,
         f10parlos_ext = b_cbclextstd_Intercept + b_cbclextstd_scalef10_neglos_c) %>% 
  pivot_longer(cols = f1parpsych_int:f10parlos_ext,
               names_to = c("ela", "behav"),
               names_sep = "_",
               values_to = "zscore",
               names_transform = list(ela = as.factor,
                                      behav = as.factor)) %>% 
  mutate(ela = case_when(
    ela == "f1parpsych" ~ "caregiver psychopathology",
    ela == "f2ses" ~ "socioeconomic disadvantage / lack of neighborhood safety",
    ela == "f3p2nosupp" ~ "lack of caregiver support (secondary)",
    ela == "f4p1nosupp" ~ "lack of caregiver support (primary)",
    ela == "f5famconf" ~ "family conflict (child report)",
    ela == "f6psusep" ~ "caregiver substance use / biological separation",
    ela == "f7angarg" ~ "family anger/arguments (caregiver report)",
    ela == "f8famagg" ~ "family verbal/physical aggression (caregiver report)",
    ela == "f9trauma" ~ "physical trauma",
    ela == "f10parlos" ~ "caregiver lack of supervision"
  ),
    behav = case_when(
      behav == "int" ~ "A. internalizing problems",
      behav == "ext" ~ "B. externalizing problems",
      behav == "cf" ~ "C. cognitive flexibility",
      behav == "ic" ~ "D. inhibitory control"
    )) %>% 
  as.data.frame()

post_cbcl_long
```
b) plot data
```{r}
png(file = "../plots/child-behav-cbcl_mvmlm-post_train.png", width = 1100, height = 550)

ggplot(data = post_cbcl_long, aes(x = ela, y = zscore, fill = ela)) +
  stat_halfeye(.width = c(.95, .80), position = "dodge") +
  facet_grid(~behav, scales = "free") +
  geom_hline(aes(yintercept = 0), linetype = "dashed", color = "darkgray") +
  labs(y = "\nassociations between early-life adversity and child behavior\n(standardized estimates)",
       x = "") +
  theme_classic() +
  coord_flip() +
  theme(axis.title = element_text(size = 22),
        axis.text = element_text(size = 18, color = "black"),
        strip.text = element_text(size = 20), # facet title text  
        legend.position = "none")

dev.off()
```

`Cognitive Control`

Get variable names:
```{r}
get_variables(m_cc_tr)
```
 

a) prepare data for plotting
```{r}
# extract posteriors from mvmlm model for int/ext problems
post_cc_long <- m_cc_tr %>% 
  spread_draws(b_cardsortstd_Intercept, b_cardsortstd_scalef1_parpsych_c,
               b_cardsortstd_scalef2_ses_adi_ncrime_c, b_cardsortstd_scalef3_p2lacksupp_c, 
               b_cardsortstd_scalef4_p1lacksupp_c, b_cardsortstd_scalef5_famconf_cr_c,
               b_cardsortstd_scalef6_parsu_biosep_c, b_cardsortstd_scalef7_angarg_pr_c, 
               b_cardsortstd_scalef8_physemoagg_pr_c, b_cardsortstd_scalef9_trauma_c,
               b_cardsortstd_scalef10_neglos_c,
               b_flankerstd_Intercept, b_flankerstd_scalef1_parpsych_c,
               b_flankerstd_scalef2_ses_adi_ncrime_c, b_flankerstd_scalef3_p2lacksupp_c, 
               b_flankerstd_scalef4_p1lacksupp_c, b_flankerstd_scalef5_famconf_cr_c,
               b_flankerstd_scalef6_parsu_biosep_c, b_flankerstd_scalef7_angarg_pr_c, 
               b_flankerstd_scalef8_physemoagg_pr_c, b_flankerstd_scalef9_trauma_c,
               b_flankerstd_scalef10_neglos_c) %>% 
  mutate(f1parpsych_cf = b_cardsortstd_Intercept + b_cardsortstd_scalef1_parpsych_c,
         f2ses_cf = b_cardsortstd_Intercept + b_cardsortstd_scalef2_ses_adi_ncrime_c,
         f3p2nosupp_cf = b_cardsortstd_Intercept + b_cardsortstd_scalef3_p2lacksupp_c,
         f4p1nosupp_cf = b_cardsortstd_Intercept + b_cardsortstd_scalef4_p1lacksupp_c,
         f5famconf_cf = b_cardsortstd_Intercept + b_cardsortstd_scalef5_famconf_cr_c,
         f6psusep_cf = b_cardsortstd_Intercept + b_cardsortstd_scalef6_parsu_biosep_c,
         f7angarg_cf = b_cardsortstd_Intercept + b_cardsortstd_scalef7_angarg_pr_c,
         f8famagg_cf = b_cardsortstd_Intercept + b_cardsortstd_scalef8_physemoagg_pr_c,
         f9trauma_cf = b_cardsortstd_Intercept + b_cardsortstd_scalef9_trauma_c,
         f10parlos_cf = b_cardsortstd_Intercept + b_cardsortstd_scalef10_neglos_c,
         f1parpsych_ic = b_flankerstd_Intercept + b_flankerstd_scalef1_parpsych_c,
         f2ses_ic = b_flankerstd_Intercept + b_flankerstd_scalef2_ses_adi_ncrime_c,
         f3p2nosupp_ic = b_flankerstd_Intercept + b_flankerstd_scalef3_p2lacksupp_c,
         f4p1nosupp_ic = b_flankerstd_Intercept + b_flankerstd_scalef4_p1lacksupp_c,
         f5famconf_ic = b_flankerstd_Intercept + b_flankerstd_scalef5_famconf_cr_c,
         f6psusep_ic = b_flankerstd_Intercept + b_flankerstd_scalef6_parsu_biosep_c,
         f7angarg_ic = b_flankerstd_Intercept + b_flankerstd_scalef7_angarg_pr_c,
         f8famagg_ic = b_flankerstd_Intercept + b_flankerstd_scalef8_physemoagg_pr_c,
         f9trauma_ic = b_flankerstd_Intercept + b_flankerstd_scalef9_trauma_c,
         f10parlos_ic = b_flankerstd_Intercept + b_flankerstd_scalef10_neglos_c) %>% 
  pivot_longer(cols = f1parpsych_cf:f10parlos_ic,
               names_to = c("ela", "behav"),
               names_sep = "_",
               values_to = "zscore",
               names_transform = list(ela = as.factor,
                                      behav = as.factor)) %>% 
  mutate(ela = case_when(
    ela == "f1parpsych" ~ "caregiver psychopathology",
    ela == "f2ses" ~ "socioeconomic disadvantage / lack of neighborhood safety",
    ela == "f3p2nosupp" ~ "lack of caregiver support (secondary)",
    ela == "f4p1nosupp" ~ "lack of caregiver support (primary)",
    ela == "f5famconf" ~ "family conflict (caregiver report)",
    ela == "f6psusep" ~ "caregiver substance use / biological separation",
    ela == "f7angarg" ~ "family anger/arguments (caregiver report)",
    ela == "f8famagg" ~ "family verbal/physical aggression (caregiver report)",
    ela == "f9trauma" ~ "physical trauma",
    ela == "f10parlos" ~ "caregiver lack of supervision"
  ),
    behav = case_when(
      behav == "int" ~ "A. internalizing problems",
      behav == "ext" ~ "B. externalizing problems",
      behav == "cf" ~ "C. cognitive flexibility",
      behav == "ic" ~ "D. inhibitory control"
    )) %>% 
  as.data.frame()

post_cc_long
```

b) plot data
```{r}
png(file = "../plots/child-behav-cogctrl_mvmlm-post_train.png", width = 1100, height = 550)

ggplot(data = post_cc_long, aes(x = ela, y = zscore, fill = ela)) +
  stat_halfeye(.width = c(.95, .80), position = "dodge") +
  facet_grid(~behav, scales = "free") +
  geom_hline(aes(yintercept = 0), linetype = "dashed", color = "gray30") +
  labs(y = "\nassociations between early-life adversity and child behavior\n(standardized estimates)",
       x = "") +
  theme_classic() +
  coord_flip() +
  theme(axis.title = element_text(size = 22),
        axis.text = element_text(size = 18, color = "black"),
        strip.text = element_text(size = 20), # facet title text 
        legend.position = "none")

dev.off()
```


# Scatterplot: spearman correlations between actual and predicted child behavior scores

a) prepare data for plotting
```{r}
df_p <- df_valid %>% 
  select(subjectkey, ends_with("_std"), starts_with("pred_")) %>% 
  rename(actual_int = cbclint_std,
         actual_ext = cbclext_std,
         actual_cf = cardsort_std,
         actual_ic = flanker_std) %>% 
  pivot_longer(cols = !subjectkey,
               names_to = c(".value", "behav"),
               names_pattern = "(.*)_(.*)") %>% 
  mutate(behav = as.factor(case_when(
    behav == "int" ~ "A. internalizing problems",
    behav == "ext" ~ "B. externalizing problems",
    behav == "cf" ~ "C. cognitive flexibility",
    behav == "ic" ~ "D. inhibitory control"
  )))
```

b) plot data
```{r}
png(file = "../plots/child-behav_pred-actual-corr.png", width = 700, height = 500)

ggplot(data = df_p, aes(x = pred, y = actual)) +
  geom_jitter(aes(color = behav, fill = behav), height = 0.25, alpha = 0.25, size = 2) +
  geom_smooth(method = "lm", color = "gray45", fill = "darkgray") +
  facet_wrap(~behav, scales = "free") +
  labs(x = "\npredicted child behavior (z-scores)",
       y = "actual child behavior (z-scores)\n") +
  theme_classic() +
  theme(axis.title = element_text(size = 22),
        axis.text = element_text(size = 18, color = "black"),
        strip.text = element_text(size = 20), # facet title text 
        legend.position = "none")

dev.off()
```

